---
title: "Analysis of Swaptions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
source(here::here("R/DataIngestion.R"))
```

## High level analysis of data from DTCC

```{r Download of data from DTCC}
start_date <- "2022-03-01"
end.date <- "2022-03-31"
bizdays <- bizdays::bizseq(start_date, end.date, "weekends")
data <- DownloadFromDTCC(bizdays)
```

```{r Filter on Swaptions}
swaptions <- data |> 
  dplyr::filter(grepl("Swaption", `Product ID`))

n.total <- nrow(swaptions)
```

The number of swaptions contracts traded in March was `r nrow(swaptions)` representing `r nrow(swaptions)/nrow(data)` of all the trades reported to DTCC.

The following table distinguishese between NEW, CANCEL and CORRECT trades

```{r Table on type of Actions}
swaptions |> 
  dplyr::count(Action)
```

Let's focus only on the new ones and see the different trade types

```{r Table on Transaction Type}
swaptions <- swaptions |> 
  dplyr::filter(grepl("NEW", Action))

swaptions |> 
  dplyr::count(`Transaction Type`)
```

```{r Filter on Transaction Type, echo=FALSE}
swaptions <- swaptions |> 
  dplyr::filter(grepl("Trade", `Transaction Type`))
```


Hence, out of `r n.total` contracts, only `r nrow(swaptions)` are actual new trades. 
We can now analyse the currencies of the new trades:

```{r Table on Currency}
swaptions |> 
  dplyr::count(`Notional Currency 1`) |> 
  dplyr::arrange(dplyr::desc(n))
```

```{r Filter on USD, echo=FALSE}
swaptions <- swaptions |> 
  dplyr::filter(grepl("USD", `Notional Currency 1`))
```


## Analysis uf USD Swaptions

Let's first analyse the style of the American USD swaptions traded in March:

```{r Option Style}
swaptions |> 
  dplyr::count(`Option Style`)
```
The vast majority is European with a few Bermudan ones.

The type of the swaption (Call/Put) should be in the field *Option Type* but that field is intelligible as it is reported:

```{r Option Type}
swaptions |> 
  dplyr::count(`Option Type`)
```

**We can possibly try and contact DTCC to get the information on what these acronyms mean?**

Let's now look at the maturities and tenors using the agreed convention

```{r Mapping on time dimensions}
trades.cube <- swaptions |> 
  dplyr::mutate(
    Maturity = round(as.numeric((`Effective Date` - `Event Timestamp`)/365),0),
    Tenor = round(as.numeric((`Expiration Date` - `Effective Date`)/365),0)) |>
  dplyr::group_by(Maturity, Tenor) |> 
  dplyr::summarise(Notional = sum(`Notional Amount 1`),
                   Trades = dplyr::n())  

trades.cube |> 
  ggplot(aes(x = Maturity, y = Tenor)) + 
  geom_tile(aes(fill = Trades), colour = "white")
```
Two considerations:

- a 120+ years Maturity contract looks like an error/outlier we can remove
- a negative maturity swaption doesn't look reasonable
- all the contracts with Maturity = 0 have a maturity which is less than 6 months

Let's remove the outlier and zoom on the 1-20 years maturity range. The number in the tile represents the number of trades for that specific pair of maturity and tenor.

```{r Filtered mapping on time dimensions}
trades.cube |> 
  dplyr::filter(Maturity <= 20, Maturity >= 0, Tenor >= 0) |> 
  ggplot(aes(x = Maturity, y = Tenor)) + 
  geom_tile(aes(fill = Trades), colour = "white") +
  geom_text(aes(label = Trades), size = 2, colour = "white") + 
  theme_minimal() + 
  theme(legend.position="none")
```

We can notice some clusters of trades, but we can further focus on the points on the cube that have at least 20 trades performed in March.

```{r Filtered mapping on time dimensions more 20 trades}
trades.cube |> 
  dplyr::filter(Maturity <= 20, Maturity >= 0, Tenor >= 0, Trades > 20) |> 
  ggplot(aes(x = Maturity, y = Tenor)) + 
  geom_tile(aes(fill = Trades), colour = "white") +
  geom_text(aes(label = Trades), size = 2, colour = "white") + 
  theme_minimal() + 
  theme(legend.position="none")
```
So most of the trades:

  - Have a maturity lower than 5 years, 7 or 10 years
  - Have tenors mostly below 10 years or 20 and 30 years
